---
title: "Week 3 Exercises"
author: "HD Sheets"
date: "10/1/2024"
output:
  html_document:
    df_print: paged
---

Please complete all exercises below. You may use any library that we have covered in class UP TO THIS POINT. 

For each function,  show that it works, by using the provided data as a test
and by feeding in some test data that you create to test your function

Add comments to your function to explain what each line is doing

1.) Write a function that takes in a string with a person's name in the form

"Sheets, Dave"

and returns a string of the form

"Dave Sheets"

Note:

-assume no middle initial ever
-remove the comma
-be sure there is white space between the first and last name

You will probably want to use stringr

```{R}
library(stringr)
name_in="Speller, Fred"

reorder_name<-function(last_first){
  # Use regex to separate last and first names, and remove the comma
  split_name <- str_split(last_first, ",\\s*")[[1]]
  
  # Rearrange to "First Last" format
  formatted_name <- str_c(split_name[2], split_name[1], sep = " ")
  return(formatted_name)
}

reorder_name(name_in)

```

2.) Write a function that takes in a string of values x, and returns a data frame
with three columns,  x, x^2 and the square root of x

```{R}
x=c(1,3,5,7,9,11,13)

powers_df<-function(x)
{
  # Create a data frame with columns: x, x^2, and sqrt(x)
   df <- data.frame(
    x = x,
    x_squared = x^2,
    x_sqrt = sqrt(x)
  )
  return(df)
}
  
powers_df(x)
```


3.) Write in a function that takes in a value x and returns
    
    y= 0.3x if x<0
    y=0.5x if x>=0
    
    This is a variant on a relu function as used in some neural networks.
```{r}
calculate_y <- function(x) {
  # if x is greater than 0 y = 0.3x
  if (x < 0) {
    y <- 0.3 * x
  } else {
    # else  y = 0.5x
    y <- 0.5 * x
  }
  return(y)
}

calculate_y(-5)
calculate_y(5)  
```

4.) Write a function that takes in a value x and returns the first power of two
greater than x  (use a While loop)
```{r}

first_power_of_two <- function(x) {
  # Start with the first power of two
  power <- 1
  
  # Keep doubling until power is greater than x
  while (power <= x) {
    power <- power * 2
  }
  
  return(power)
}

first_power_of_two(5) 
first_power_of_two(15)
```


5) Two Sum - Write a function named two_sum()

Given a vector of integers nums and an integer target, return indices of the two numbers such that they add up to target.

You may assume that each input would have exactly one solution, and you may not use the same element twice.

You can return the answer in any order.

Example 1:

Input: nums = [2,7,11,15], target = 9
Output: [0,1]
Explanation: Because nums[0] + nums[1] == 9, we return [0, 1].
Example 2:

Input: nums = [3,2,4], target = 6
Output: [1,2]
Example 3:

Input: nums = [3,3], target = 6
Output: [0,1]

Constraints:

2 <= nums.length <= 104
–109 <= nums[i] <= 109
–109 <= target <= 109
Only one valid answer exists.
 

*Note: For the first problem I want you to use a brute force approach (loop inside a loop)* 

*The brute force approach is simple. Loop through each element x and find if there is another value that equals to target – x*

*Use the function seq_along to iterate*

```{r}
two_sum <- function(nums_vector, target) {
  # Initialize an empty list to store results
  result <- list()
  
  # Outer loop over each element
  for (i in seq_along(nums_vector)) {
    
    
    # Inner loop to check remaining elements for a match
    for (j in (i + 1):length(nums_vector)) {
      
      
      # Check if both elements are non-NA and their sum equals the target
     if (!is.na(nums_vector[i]) && !is.na(nums_vector[j]) && 
          nums_vector[i] + nums_vector[j] == target) {
        
        
        # Add indices as a pair to result
        result <- append(result, list(c(i, j)))
      }
    }
  }
  
  # Return the list of index pairs
  return(result)
}




# Test code
nums_vector <- c(5,7,12,34,6,10,8,9)
target <- 13
 
z=two_sum(nums_vector,target)
print(z)
#expected answers
#[1] 1 7
#[1] 2 5
#[1] 5 2

```

6.) Write one piece of code that will use a regex command to extract a phone
number written in the form

      456-123-2329
      
The sentences to use are located below

use the str_extract function from stringr

use the same regex search pattern from each

-What does \\d match to?  or alternatively [[:digit:]]

-How do you specify a specific number of repeated characters

```{R}
a="Please call me at 456-123-2329, asap"
b="Hey, we have a code 234 on machine a-234-12, call me at 678-321-9876"
c="On 12-23-2022, Joe over at 122 Turnpike, dialled 912-835-4756, tell me by 9:02 pm Wed"

phone_pattern <- "\\b\\d{3}-\\d{3}-\\d{4}\\b"

# Extract the phone numbers
phone_a <- str_extract(a, phone_pattern)
phone_b <- str_extract(b, phone_pattern)
phone_c <- str_extract(c, phone_pattern)

# Print the results
print(phone_a) 
print(phone_b)
print(phone_c) 
```

7.)  For lines below,  extract the domains (ie the part of the address after @)

```{R}
d="jimmy.halibut@gmail.com"
e="His address is:  c.brown@hopeles.org, do write him"
f="h.potter@hogwarts.edu is bouncing back on me, I wonder why?"

# Define the regex pattern to match the domain part after @
domain_pattern <- "@([A-Za-z0-9.-]+\\.[A-Za-z]{2,})"

# Extract the domains
domain_d <- str_extract(d, domain_pattern) %>% str_remove("^@")
domain_e <- str_extract(e, domain_pattern) %>% str_remove("^@")
domain_f <- str_extract(f, domain_pattern) %>% str_remove("^@")

# Print the results
print(domain_d)
print(domain_e) 
print(domain_f)  
```

